On Math

### 2025-07-02 cards/paper.tex
quotient rule for differentiation
product rule for differentiation
push "integrals of trigonometry functions"
function discontinuity

@@

\subsection{Chain rule for differentiation}
Derivative of a composite function $f(g(x))$ is given by:
\begin{equation}
  \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} = \frac{df}{dx}
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
    f(x) = (3x^4 + x^3)^5 \\
    \frac{df}{dx} = ? \\
    f = g^5 \\
    g = 3x^4 + x^3 \\
    \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} \\
    \frac{df}{dg} = 5g^4 \\
    \frac{dg}{dx} = 12x^3 + 3x^2 \\
    \frac{df}{dx} = 5g^4 * (12x^3 + 3x^2) = 5(3x^4 + x^3)^4(12x^3 + 3x^2)
  \end{gathered}
\end{equation}
The chain rule also can be expressed using primes instead of Leibniz notation
\begin{equation}
  f(g(x))' = f'(g(x)) * g'(x)
\end{equation}
\subsection{Differentiation of an exponent function}
\begin{equation}
  \frac{d}{dx} (a^x) = a^x\ln a
\end{equation}
with special case:
\begin{equation}
  \frac{d}{dx} (e^x) = e^x\ln e = e^x * 1 = e^x
\end{equation}
\subsection{Chain rule for exponent function}
\begin{equation}
  \begin{gathered}
    y = e^{3x} \\
    y = e^u \\
    u = 3x \\
    \frac{dy}{dx} = \frac{dy}{du} * \frac{du}{dx} \\
    \frac{d}{du} e^u = e^u \\
    \frac{d}{dx} 3x = 3 \\
    \frac{dy}{dx} = e^u * 3 = 3e^u = 3e^{3x}
  \end{gathered}
\end{equation}

@@

>>>>

\subsection{Function continuity}
If a function $f(x)$ is \textbf{continuous}, we can draw $y = f(x)$ without
taking our pen off the paper. A function $f(x)$ is continuous at the point
$x = c$ if:

\begin{equation}
  \lim_{x \to c}f(x) = f(c)
\end{equation}

To test for continuity of $f(x)$ we need to check three conditions:

1. does $f(c)$ exists?
We must check that $f(x)$ is defined at $x=c$

2. does $\lim_{x \to c}f(x)$ exists?
We must have to check that $\lim_{x \to c^-}f(x) = \lim{x \to c^+}f(x)$

3. does $\lim{x \to c}f(x) = f(c)$

If all of the above is yes, then the function \textbf{is continuous} at $x = c$.
Otherwise we say there's \textbf{discontinuity} at $x = c$.

\begin{tikzpicture}[scale=1.2]
    % Draw axes
    \draw[->] (-0.5,0) -- (5,0) node[right] {$x$};
    \draw[->] (0,-0.5) -- (0,4) node[above] {$y$};

    % Draw the function curve (looks like a logarithmic or similar curve)
    % Left part approaching the discontinuity
    \draw[thick] plot[domain=0.1:5, samples=100] (\x, {1.5 + 0.8*ln(\x)});

    % Mark the discontinuity point
    \coordinate (c) at (3,0);
    \coordinate (leftlimit) at (3,2.38);

    % Draw vertical dashed line at x = c
    \draw[dashed] (3,0) -- (3,3.5);

    % Draw horizontal dashed line showing limit
    \draw[dashed] (0,2.4) -- (3,2.4);

    % Mark points
    \fill (leftlimit) circle (2pt);

    % Add labels
    \node[below] at (3,0) {$c$};
    \node[left] at (0,2.4) {$\lim_{x \to c}f(x)$};
    \node[right] at (4.5,2.5) {$y = f(x)$};

    % Add limit annotations
    \node at (2,0.5) {$x \to c^-$};
    \node at (4,0.5) {$x \to c^+$};

    % Add arrows showing approach for x
    \draw[->] (2.5,0.3) -- (2.8,0.3);
    \draw[->] (3.5,0.3) -- (3.2,0.3);

    % Add arrows showing approach for y
    \node at (0.5,2.8) {$\downarrow f(x)$};
    \node at (0.5,2.0) {$\uparrow f(x)$};

\end{tikzpicture}

@@

\subsection{Chain rule for differentiation}
Derivative of a composite function $f(g(x))$ is given by:
\begin{equation}
  \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} = \frac{df}{dx}
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
    f(x) = (3x^4 + x^3)^5 \\
    \frac{df}{dx} = ? \\
    f = g^5 \\
    g = 3x^4 + x^3 \\
    \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} \\
    \frac{df}{dg} = 5g^4 \\
    \frac{dg}{dx} = 12x^3 + 3x^2 \\
    \frac{df}{dx} = 5g^4 * (12x^3 + 3x^2) = 5(3x^4 + x^3)^4(12x^3 + 3x^2)
  \end{gathered}
\end{equation}
The chain rule also can be expressed using primes instead of Leibniz notation
\begin{equation}
  f(g(x))' = f'(g(x)) * g'(x)
\end{equation}
\subsection{Product rule for differentiation}
for $f(x) = u(x)*v(x)$ differentiation is:
\begin{equation}
  \begin{gathered}
    f'(x) = u'(x)*v(x) + u(x)*v'(x) \\
    \text{Using Leibnitz:} \\
    \frac{dy}{dx} = \frac{du}{dx}*v + \frac{dv}{dx}*u
  \end{gathered}
\end{equation}
\subsection{Quotient rule for differentiation}
for $f(x) = \frac{u(x)}{v(x)}$ differentiation is:
\begin{equation}
  \begin{gathered}
    f'(x) = \frac{u'(x)*v(x) - u(x)*v'(x)}{[v(x)]^2} \\
    \text{Using Leibnitz:} \\
    \frac{d}{dx}(\frac{u}{v}) = \frac{v*\frac{du}{dx} - u*\frac{dv}{dx}}{v^2}
  \end{gathered}
\end{equation}
\subsection{Differentiation of an exponent function}
\begin{equation}
  \frac{d}{dx} (a^x) = a^x\ln a
\end{equation}
with special case:
\begin{equation}
  \frac{d}{dx} (e^x) = e^x\ln e = e^x * 1 = e^x
\end{equation}
\subsection{Chain rule for exponent function}
\begin{equation}
  \begin{gathered}
    y = e^{3x} \\
    y = e^u \\
    u = 3x \\
    \frac{dy}{dx} = \frac{dy}{du} * \frac{du}{dx} \\
    \frac{d}{du} e^u = e^u \\
    \frac{d}{dx} 3x = 3 \\
    \frac{dy}{dx} = e^u * 3 = 3e^u = 3e^{3x}
  \end{gathered}
\end{equation}

@@

\subsection{Integrals of trigonometry functions}
\begin{equation}
  \begin{gathered}
    \int \sin x dx = -\cos x + C \\
    \int \cos x dx = \sin x + C \\
    \int \sec^2 x dx = \tan x + C \\
    \int \sec x \tan x dx = \sec x + C \\
    \int \csc x \cot x dx = -\csc x + C \\
    \int \csc^2 x dx = -\cot x + C
  \end{gathered}
\end{equation}
